\thetitle

### 2025-06-28 workshop/PT.tex
minor changes

\subsection{Overview of PT}
Suppose that we would like to estimate integrals involving $\pi$, such as 
$\int f(x) \pi(x) \, \dee x$. 
These integrals may be multivariate and even include combinations of 
continuous and discrete variables (where sums replace integrals in the discrete 
case).
One method is to obtain samples from $\pi$ to approximate such integrals.
Often, the distribution 
$\pi$ can be challenging for traditional MCMC methods---such 
as Metropolis-Hastings, slice sampling, and Hamiltonian Monte Carlo---
because of its structure.
For example, in a bimodal example such as the one illustrated in 
\cref{fig:bimodal},
traditional methods might remain in one of the two modes for an extremely 
long period of time.

>>>>

\subsection{Overview of PT}
Suppose that we would like to estimate integrals involving $\pi$, such as 
$\int f(x) \pi(x) \, \dee x$. 
These integrals may be multivariate and even include combinations of 
continuous and discrete variables (where sums replace integrals in the discrete 
case).
One method is to obtain samples from $\pi$ to approximate such integrals.
Often, the distribution 
$\pi$ can be challenging for traditional MCMC methods---such 
as Metropolis--Hastings, slice sampling, and Hamiltonian Monte Carlo---
because of its structure.
For example, in a bimodal example such as the one illustrated in 
\cref{fig:bimodal},
traditional methods might remain in one of the two modes for an extremely 
long period of time.

### 2025-06-28 workshop/nutshell.tex
minor changes

\section{Parallel Tempering in a nutshell}

\begin{figure}[t]
  \centering
    \includegraphics[width=0.5\linewidth]{../img/bimodal.pdf}
    \caption{A bimodal distribution. Blue lines display output 
    from 1,000 iterations of a random walk Metropolis-Hastings sampler.}
    \label{fig:bimodal}
\end{figure}

Consider the problem of sampling from a probability density (or mass) function
$\pi(x)$, henceforth called the \emph{target}. Solving this task is often 
challenging, as the 
distribution $\pi$ can exhibit features which traditional MCMC methods---from
random walk Metropolis-Hastings to Hamiltonian Monte Carlo---struggle to
describe accurately. For example, in bimodal distributions such as the one 
illustrated in \cref{fig:bimodal}, traditional methods might remain stuck in one 
of the two modes for an extremely long period of time.

\begin{figure}[t]
  \centering
    \begin{minipage}{0.333\linewidth}
      \centering
      \includegraphics[width=\linewidth]{../img/heatmap_path_1.pdf}
%      \caption*{$\pi_1$}
    \end{minipage}%
    \begin{minipage}{0.333\linewidth}
      \centering
      \includegraphics[width=\linewidth]{../img/heatmap_path_2.pdf}
%      \caption*{$\pi_2$}
    \end{minipage}%
    \begin{minipage}{0.333\linewidth}
      \centering
      \includegraphics[width=\linewidth]{../img/heatmap_path_3.pdf}
%      \caption*{$\pi_3$}
    \end{minipage}\\
    \begin{minipage}{0.333\linewidth}
      \centering
      \includegraphics[width=\linewidth]{../img/heatmap_path_4.pdf}
%      \caption*{$\pi_4$}
    \end{minipage}%
    \begin{minipage}{0.333\linewidth}
      \centering
      \includegraphics[width=\linewidth]{../img/heatmap_path_5.pdf}
%      \caption*{$\pi_5$}
    \end{minipage}%
    \begin{minipage}{0.333\linewidth}
      \centering
      \includegraphics[width=\linewidth]{../img/heatmap_path_6.pdf}
%      \caption*{$\pi_6$}
    \end{minipage}
    \caption{Heatmaps of six distributions lying on an annealing path, 
    starting from a unimodal reference (top-left), and ending at a target
    distribution from \cref{fig:bimodal}.}
    \label{fig:path}
\end{figure}

>>>>

\section{Parallel tempering in a nutshell}

\begin{figure}[t]
  \centering
    \includegraphics[width=0.5\linewidth]{../img/bimodal.pdf}
    \caption{A bimodal distribution. Blue lines display output 
    from 1,000 iterations of a random walk Metropolis--Hastings sampler.}
    \label{fig:bimodal}
\end{figure}

Consider the problem of sampling from a probability density (or mass) function
$\pi(x)$, henceforth called the \emph{target}. Solving this task is often 
challenging, as the 
distribution $\pi$ can exhibit features which traditional MCMC methods---from
random walk Metropolis--Hastings to Hamiltonian Monte Carlo---struggle to
describe accurately. For example, in bimodal distributions such as the one 
illustrated in \cref{fig:bimodal}, traditional methods might remain stuck in one 
of the two modes for an extremely long period of time.

\begin{figure}[t]
  \centering
    \begin{minipage}{0.333\linewidth}
      \centering
      \includegraphics[width=\linewidth]{../img/heatmap_path_1.pdf}
%      \caption*{$\pi_1$}
    \end{minipage}%
    \begin{minipage}{0.333\linewidth}
      \centering
      \includegraphics[width=\linewidth]{../img/heatmap_path_2.pdf}
%      \caption*{$\pi_2$}
    \end{minipage}%
    \begin{minipage}{0.333\linewidth}
      \centering
      \includegraphics[width=\linewidth]{../img/heatmap_path_3.pdf}
%      \caption*{$\pi_3$}
    \end{minipage}\\
    \begin{minipage}{0.333\linewidth}
      \centering
      \includegraphics[width=\linewidth]{../img/heatmap_path_4.pdf}
%      \caption*{$\pi_4$}
    \end{minipage}%
    \begin{minipage}{0.333\linewidth}
      \centering
      \includegraphics[width=\linewidth]{../img/heatmap_path_5.pdf}
%      \caption*{$\pi_5$}
    \end{minipage}%
    \begin{minipage}{0.333\linewidth}
      \centering
      \includegraphics[width=\linewidth]{../img/heatmap_path_6.pdf}
%      \caption*{$\pi_6$}
    \end{minipage}
    \caption{Heatmaps of six distributions lying on an annealing path, 
    starting from a unimodal reference (top-left), and ending at the target
    distribution from \cref{fig:bimodal}.}
    \label{fig:path}
\end{figure}
